---
output: pdf_document
editor_options: 
  chunk_output_type: inline
author: "Zach Bogart"
title: "Advent of Code 2020"
date: "`r Sys.Date()`"
---

# Day 10: Adapter Array

[Click for Problem Statement](https://adventofcode.com/2020/day/10)

---

```{r include=FALSE}
library(tidyverse)
```

```{r}
testing1 = tibble(x = as.numeric(readLines("inputs/10-test1.txt")))
testing2 = tibble(x = as.numeric(readLines("inputs/10-test2.txt")))
input = tibble(x = as.numeric(read_lines("inputs/10-input.txt")))
```

\newpage

# Part 1

## Testing

```{r}
t2_p1 = testing2 %>% 
  add_row(x = 0) %>% # charging outlet
  add_row(x = max(.$x, na.rm=TRUE) + 3) %>%  # end adapter
  arrange(x) %>% 
  mutate(diff = x - lag(x)) %>% 
  mutate(j1 = diff == 1,
         j3 = diff == 3) 

t2_p1 %>% 
  summarise(total_j1 = sum(j1, na.rm = TRUE),
            total_j3 = sum(j3, na.rm = TRUE),
            answer = total_j1 * total_j3)
```

## The Real Thing

```{r}
part1 = input %>% 
  add_row(x = 0) %>% # charging outlet
  add_row(x = max(.$x, na.rm=TRUE) + 3) %>%  # end adapter
  arrange(x) %>% 
  mutate(diff = x - lag(x)) %>% 
  mutate(j1 = diff == 1,
         j3 = diff == 3)

part1 %>% 
  summarise(total_j1 = sum(j1, na.rm = TRUE),
            total_j3 = sum(j3, na.rm = TRUE),
            answer = total_j1 * total_j3)
```

\newpage

# Part 2

## Testing

Doing some working out, seems to be that each additional adapter in a string of 1-jolts gives a multiple of three more options.

- need to find strings of ones, count size of string, and multiply results based on rules

- One 1-jolt provides only one option
- Two 1-jolts provide only two options
- For `n >= 3` 1-jolts, they provide `3(n-2) + 1` options

```{r}
t2_p2 = t2_p1 %>% 
  filter(!is.na(diff)) %>% 
  mutate(one_string = cumsum(diff == 3)) %>% 
  filter(diff != 3) %>% 
  group_by(one_string) %>% 
  summarise(options = ifelse(n() == 1, 1, ifelse(n() == 2, 2, 3*(n()-2) + 1)))

t2_p2 %>% 
  summarise(answer = prod(options))
```

## The Real Thing

```{r}
options(scipen=30)

part2 = part1 %>% 
  filter(!is.na(diff)) %>% 
  mutate(one_string = cumsum(diff == 3)) %>% 
  filter(diff != 3) %>% 
  group_by(one_string) %>% 
  summarise(options = ifelse(n() == 1, 1, ifelse(n() == 2, 2, 3*(n()-2) + 1)))

part2 %>% 
  summarise(answer = format(prod(options), big.mark = ","))
```


