---
output: html_document
editor_options: 
  chunk_output_type: inline
---

# Day 3: Toboggan Trajectory

https://adventofcode.com/2020/day/3

```{r include=FALSE}
library(tidyverse)
```


## Part 1

- Input, going **right 3 down 1** every row
- Given patterns repeat

```{r message=FALSE}
raw = read_delim("03-input.txt", delim="\n", col_names = "text")
```

How long is every string?

```{r}
raw %>% 
  rowwise() %>% 
  mutate(length = nchar(text))
```
Every string is 31 characters long. There are 323 rows.

- For n rows, need each string to be 3(n-1)+1 characters long, or 3n-2
- Since there are 323 rows, n = 323
- So, every string needs to be 967 characters long

Can repeat each string 32 times to allow for enough input

```{r}
repeat_1 = ceiling((3 * nrow(raw) - 2) / 31)
```

```{r}
input = raw %>% 
  mutate(text = strrep(text, repeat_1),
         row = row_number(),
         index_value = 3 * row - 2,
         hit_a_tree = (str_sub(text, index_value, index_value) == "#"))

sum(input$hit_a_tree)
```

We hit 171 trees.

## Part 2

- add more slopes this time!

- Right 1, down 1.
- Right 3, down 1. (This is the slope you already checked.)
- Right 5, down 1.
- Right 7, down 1.
- Right 1, down 2.

Should all be straight forward except a little filtering for the down 2 example. Will need to repeat the string a few more times.

- 7(n-1)+1 or 7n-6
- Every string needs to be 2255 character long
- Need to repeat each string 73 times (each string 31 chars long)

```{r}
repeat_2 = ceiling((7 * nrow(raw) - 6) / 31)
```

```{r}
slope_1 = raw %>% 
  mutate(text = strrep(text, repeat_2),
         row = row_number(),
         
         right_1 = 1 * (row-1) + 1,
         right_3 = 3 * (row-1) + 1,
         right_5 = 5 * (row-1) + 1,
         right_7 = 7 * (row-1) + 1,
         
         hit_1 = (str_sub(text, right_1, right_1) == "#"),
         hit_3 = (str_sub(text, right_3, right_3) == "#"),
         hit_5 = (str_sub(text, right_5, right_5) == "#"),
         hit_7 = (str_sub(text, right_7, right_7) == "#"))

# the down 1 slopes
down_1 = slope_1 %>% 
  select(hit_1:hit_7) %>% 
  summarise_all(sum)

# the down 2 slopes
# have to make sure to filter by index correctly! Don't define row value till after!
slope_2 = raw %>% 
  filter(row_number() %% 2 == 1) %>% 
  mutate(text = strrep(text, 73),
         row = row_number()) %>% 
  mutate(right_1 = 1 * (row-1) + 1,
         hit_1 = (str_sub(text, right_1, right_1) == "#"))

down_2 = slope_2 %>% 
  select(hit_1) %>% 
  summarise(hit_1_down_2 = sum(hit_1))

results = cbind(down_1, down_2)

results
```

Above are how many trees we hit for each slope. So, their product is...

```{r}
results %>% 
  rowwise() %>% 
  prod()
```

1206576000

:)